Transition path sampling based calculations of free energies of enzymatic reactions: the case of human methionine adenosyl transferase and plasmodium vivax adenosine deaminase

### 2022-04-27 tex/manuscript.tex
added some pretty pictures

\begin{figure}
\includegraphics[scale=0.6]{figures/mat2a-reaction.png}
\caption{The reaction catalyzed by human MAT2A enzyme}
\label{fig:mat2a-reaction}
\end{figure}

@@

%------------------------------------------------------------------------------ 
\subsection{Free energy calculations}
%------------------------------------------------------------------------------ 
To obtain the free energy profile along the order parameter, we use the method 
discussed in sec. \ref{ssec:algorithm}. The order paratemer for the MAT2A system 
is defined to be $dis(OC)-dis(SC)$ and that for the $pv$ADA is defined as 
$dis(NC)-dis(OC)$. For the MAT2A system, configurations are sampled within 
the range $[-2.0,2.0]$ and this range is divided into 14 overlapping windows 
($0.09$ {\AA} overlap on either side). 
Configurations are sampled for the $pv$ADA system within the interval $[-3.0, 3.0]$ {\AA}
of the order parameter value. 
This interval is divided into 20 overlapping windows ($0.045$ {\AA} between 
neighboring windows). A reactive trajectory from the TPS ensemble is chosen to be the guide trajectory 
for sampling within each of the windows. To sample in a particular window $w$, a 
frame from this reactive trajectory is chosen such that the value of its 
order parameter falls within $w$. Using the shooting algorithm
the momenta are perturbed and dynamics is run for 10 steps forwards and backwards with a 
$1\;fs$ time step to obtain a trajectory with a time period of $20\;fs$.    
Within each window 3000 TPS trajectories are collected. The 
probability distribution function for the order parameter within window $w$ 
is then calculated as normalized histograms which represent the populations 
of the configurations within the window at equilibrium. 
Then we calculate the free energy profile in a similar fashion to the WHAM 
procedure \cite{Kumar92JComputChem13p1011} which is generally used to obtain the 
potential of mean force from umbrella sampling calculations. 
%------------------------------------------------------------------------------ 
\section{Results and discussion}
%------------------------------------------------------------------------------ 
\subsection{Human Methionine Adenosyltransferase 2A}
The binding pocket image is shown for this enzyme here
% width: 3.330, height: 1.873
The TPS ensemble was used to analyze the reaction pathway for the S$_{\text{N}}2$
mechanism and a typical trajectory is provided here
\begin{figure}
\includegraphics[scale=1.0]{figures/mat2a-diff60.pdf}
\caption{A reactive trajectory}
\label{fig:mat2a-reactive-traj}
\end{figure}
The order parameters of the system along a typical reactive trajectory is 
shown in Fig. \ref{fig:mat2a-reactive-traj}.

\begin{figure}[ht!]
\includegraphics[scale=0.25]{figures/mat2a-transition-state.png}
\end{figure}

\begin{figure}[ht!]
\includegraphics[scale=0.5]{figures/comm-60-mat2a.pdf}
\end{figure}

@@

%------------------------------------------------------------------------------ 
\subsection{Adenosine deaminase reaction coordinates and free energy}
%------------------------------------------------------------------------------ 
\begin{figure}[ht!]
\includegraphics[scale=0.15]{./figures/ada-protein.png}
\end{figure}

\begin{figure}[ht!]
\includegraphics[scale=0.5]{./figures/ada-fenergy.pdf}
\end{figure}

>>>>

\begin{figure}
\includegraphics[width=0.5\textwidth]{figures/mat2a-reaction.png}
\caption{The reaction catalyzed by human MAT2A enzyme}
\label{fig:mat2a-reaction}
\end{figure}

@@

%------------------------------------------------------------------------------ 
\subsection{Free energy calculations}
%------------------------------------------------------------------------------ 
To obtain the free energy profile along the order parameter, we use the method 
discussed in sec. \ref{ssec:algorithm}. The order paratemer for the MAT2A system 
is defined to be $dis(OC)-dis(SC)$ and that for the $pv$ADA is defined as 
$dis(NC)-dis(OC)$. For the MAT2A system, configurations are sampled within 
the range $[-2.0,2.0]$ and this range is divided into 14 overlapping windows 
($0.09$ {\AA} overlap on either side). 
Configurations are sampled for the $pv$ADA system within the interval $[-3.0, 3.0]$ {\AA}
of the order parameter value. 
This interval is divided into 20 overlapping windows ($0.045$ {\AA} between 
neighboring windows). A reactive trajectory from the TPS ensemble is chosen to be the guide trajectory 
for sampling within each of the windows. To sample in a particular window $w$, a 
frame from this reactive trajectory is chosen such that the value of its 
order parameter falls within $w$. Using the shooting algorithm
the momenta are perturbed and dynamics is run for 10 steps forwards and backwards with a 
$1\;fs$ time step to obtain a trajectory with a time period of $20\;fs$.    
Within each window 3000 TPS trajectories are collected. The 
probability distribution function for the order parameter within window $w$ 
is then calculated as normalized histograms which represent the populations 
of the configurations within the window at equilibrium. 
Then we calculate the free energy profile in a similar fashion to the WHAM 
procedure \cite{Kumar92JComputChem13p1011} which is generally used to obtain the 
potential of mean force from umbrella sampling calculations. 
%------------------------------------------------------------------------------ 
\section{Results and discussion}
%------------------------------------------------------------------------------ 
\subsection{Human Methionine Adenosyltransferase 2A}
\begin{figure}[ht!]
\includegraphics[scale=0.2]{figures/mat2a-equil.png}
\label{fig:mat2a-equil}
\caption{Equilibriated structure of MAT2A enzyme}
\end{figure}

The binding pocket image is shown for this enzyme here
% width: 3.330, height: 1.873
The TPS ensemble was used to analyze the reaction pathway for the S$_{\text{N}}2$
mechanism and a typical trajectory is provided here
\begin{figure}
\includegraphics[scale=1.0]{figures/mat2a-diff60.pdf}
\caption{A reactive trajectory}
\label{fig:mat2a-reactive-traj}
\end{figure}
The order parameters of the system along a typical reactive trajectory is 
shown in Fig. \ref{fig:mat2a-reactive-traj}.

\begin{figure}[ht!]
\includegraphics[scale=0.2]{figures/mat2a-transition-state.png}
%\includegraphics[scale=0.2]{figures/mat2a-active-site2.png}
\end{figure}

\begin{figure}[ht!]
\includegraphics[scale=1.0]{figures/comm-60-mat2a.pdf}
\end{figure}

@@

%------------------------------------------------------------------------------ 
\subsection{Adenosine deaminase reaction coordinates and free energy}
%------------------------------------------------------------------------------ 
\begin{figure}[ht!]
\centering
\includegraphics[scale=0.2]{./figures/ada-equil.png}
\caption{Equilibriated structure of ADA protein}
\end{figure}

\begin{figure}[ht!]
\includegraphics[scale=1.0]{./figures/ada-fenergy.pdf}
\label{fig:ada-fenergy}
\caption{The free energy curve for the ADA reaction}
\end{figure}
